Tail associations in ecological variables and their impact on extinction risk

### 2020-03-23 MainText.Rmd
made changes in Rmds as DAN said

<!--stochastic matrix model : age-structured-->
To investigate Q1 from the Introduction,
we considered a simple stochastic age-structured model
with two age classes, 
\begin{equation}
\left({\begin{array}{cc} E(t+1) \\ A(t+1) \end{array}}\right) = 
\left(\begin{array}{cc} 0 & f_a(t)\\ s_e(t) & 0 \end{array}\right)
\left(\begin{array}{c} E(t) \\ A(t) \end{array}\right), \label{eq_age_str}
\end{equation}
\noindent where $E(t)$ and $A(t)$ are the numbers or population densities of eggs 
and adults, respectively, at time $t$. We chose a very simple model
because our goal was to carry out an initial evaluation, only, of 
the realism of a potential mechanism. The stochastic vital 
rates $s_e(t)$ and $f_a(t)$ in Eq. \ref{eq_age_str} were selected, independently 
across times, as follows. The egg survival rate $s_e(t)$ was 
drawn from a beta distribution with parameters $\alpha=3$ and $\beta=2$. 
Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
was drawn from a gamma distribution with shape and scale parameters $k=2$
and $\theta=1$, respectively, 
and $\Omega$ is a density dependence parameter. Taking $\Omega=\infty$ made
$\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
(it was then a standard stochastic matrix model). Finite, positive values of 
$\Omega$ made the model nonlinear, with density dependence in the adult
fecundity rate. Stochastic matrix models are very commonly used in
evolutionary ecology and quantitative conservation biology, so our model
is an example of a very widely used category of models. See, for instance,
the classic books of @Caswell2000, and @MorrisDoak2002.

@@

The variables $u(t)$ and $v(t)$ are visibly asymmetrically tail associated (Fig. \ref{fig_intro_cop}),
but we have instead used the transformed variables $s_e(t)$ and $\tilde{f}_a(t)$
in our model. In what sense are $s_e(t)$ and $\tilde{f}_a(t)$ asymmetrically
tail associated? Given two boundary quantities $l_b$ and $u_b$ such
that $0 \leq l_b < u_b \leq 1$, and letting $x(t)$ be the rank of 
$u(t)$ in the set $\{u(t) | t=1,\ldots,T\}$, divided by $T+1$, and letting 
$y(t)$ be the rank of 
$v(t)$ in the set $\{v(t) | t=1,\ldots,T\}$, divided by $T+1$ (smallest elements
are here defined to have rank 1),
consider the portions of the 
lines $x+y = 2l_b$ and $x+y=2u_b$ which lie within the unit square
(Fig. \ref{PartialSpearmanFig}). 
We can then define the *partial Spearman correlation*,
\begin{equation}
\cor_{l_b,u_b}(u,v) = \frac{\sum (x(t)-\mean(x))(y(t)-\mean(y))}{(n-1)\sqrt{\var(x) \var(y)}},\label{PartialSpearmanEq}
\end{equation}
\noindent where sample means and variances in this equation
are computed using all the data but the sum is computed only
over the $t$ such that $x(t)+y(t) > 2l_b$ and $x(t)+y(t)<2u_b$. The variance, $\var$,
here and throughout, is the sample variance, with Bessel's correction (denominator $T-1$).
The partial Spearman correlation is the component of the Spearman
correlation of $u$ and $v$ which can be attributed to the points
for which the normalized ranks, $x$ and $y$, lie between the boundary lines
we defined in
the unit square (Fig. \ref{PartialSpearmanFig}). 
The quantity $\cor_{0,0.5}(u,v)-\cor_{0.5,1}(u,v)$ 
is a way to measure asymmetry of tail associations between
$u$ and $v$. And this quantity is the same as 
$\cor_{0,0.5}(s_e,\tilde{f}_a)-\cor_{0.5,1}(s_e,\tilde{f}_a)$, 
because the partial Spearman correlation, like the Spearman
correlation itself, is based on ranks. Ranks are the same
for $u(t)$ and for $s_e(t)$ because these quantities are 
related by an increasing monotonic transformation; and likewise for 
$v(t)$ and $\tilde{f}_a(t)$. Thus, in the sense captured
by the statistic $\cor_{0,0.5}-\cor_{0.5,1}$, $s_e(t)$
and $\tilde{f}_a(t)$ have the same asymmetry of tail
association as the values $u(t)$ and $v(t)$ from which they
were constructed. Rank-based approaches to understanding the nature
of association between variables have been recommended, as such
approaches are not influenced by the marginal distributions of
the quantities (e.g., @Genest2007). Marginal distributions contain no information
about association between variables, as each marginal pertains to
only one of the variables. These ideas are revisited in the Discussion.

@@

<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
To answer (Q2) of the Introduction, 
we estimated extinction risk for each metapopulation model, using each form of tail 
association for the environmental noise,
using values of $r$ that included under- and overcompensatory dynamics,
as indicated above. We chose values of $r$ within a range, the
range specified as follows. The second part of Q2 (about how density dependence mediates the
influence of asymmetric tail associations on extinction risk) was addressed by comparing
results for under- and overcompensatory model regimes.
Defining $r_{min}$ as the minimum value of $r$ 
for which $P_e$ was non-negative; and defining
$r_{bf}$ as the value of $r$ at which a bifurcation occurs, the range of
$r$ used spanned 95\% of the interval ($r_{min},r_{bf}$). For details, see Appendix 
\ref{SM-model_stability}. The range of $r$ encompassed $r_c$, so included 
a transition from under- to overcompensatory dynamics. For each value of $r$
used within the range, we calculated extinction risk as indicated above.
Dispersal parameter values $d= 0,0.1,0.2,\ldots,0.5$ were also used in different 
simulations. We plotted extinction risk as a function of $d$ for several
fixed values of $r$, some producing under- and some overcompensatory dynamics.
For $d=0$ we also plotted extinction risk as a function of $r$. All code for 
this project is available at https://github.com/sghosh89/ERC.

@@

<!--Q2 : metapopulation results-->
For our metapopulation simulations, for low values of $r$ (close to $r_{min}$), left-tail
associations of environmental noises accentuated extinction risks, and right-tail associations
mitigated them; but the reverse was true for high values of $r$ (close to
$r_{bf}$). Thus 
density dependent models can either produce similar results, regarding the influence of 
tail associations on extinction risk, as the density-independent models of @Ghosh_copula, or
opposite results, depending on the nature of density dependence. 
For each model of Table \ref{tab_model_summary}, using a value of $r$ on
the low end of the range $(r_{min},r_{bf})$, extinction risk was highest
for extreme left-tail associated noise, lowest for extreme right-tail associated
noise, and was intermediate for the moderate tail association scenarios, for
both local (Fig. \ref{fig_extrisk_allmodel_localdisp}, left panels) and global 
(Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}, left panels) dispersal, and for 
essentially all the dispersal rates, $d$, that we considered. However, 
using a value of $r$ on the high end of the range $(r_{min},r_{bf})$,
extinction risks were ordered oppositely (right panels of Figs 
\ref{fig_extrisk_allmodel_localdisp} and \ref{SM-fig_extrisk_allmodel_globaldisp}). 
These results help address Q2 from the Introduction.
Intermediate values of $r$ are shown in Figs. \ref{SM-fig_ricker_risk_local_disp_scl_1} - 
\ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} for local dispersal and 
Figs. \ref{SM-fig_ricker_risk_global_disp_scl_1} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6} for global dispersal among 
habitat patches.

Because the deterministic, one-patch versions of our models showed undercompensatory
dynamics for $r<r_c$ (and therefore for $r$ on the left end of the range $(r_{min},r_{bf})$),
and showed overcompensatory dynamics for $r>r_c$, Figs \ref{fig_extrisk_allmodel_localdisp}
and \ref{SM-fig_extrisk_allmodel_globaldisp} suggest the hypothesis that the
opposite effects of tail association on extinction risk correspond to the domains
in which models exhibit under- versus overcompensatory dynamics. This was 
approximately, but not exactly, true. Careful inspection of Figs 
\ref{SM-fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{SM-fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6} 
and the $r_c$ values of Table \ref{tab_model_summary} reveals that, for the Verhulst and Austin-Brewer models,
the switch occurs at a value of $r$ less than $r_c$. For the example
case of $d=0$, plots of extinction risk versus $r$ (Fig. \ref{fig_extrisk_d0}) seem to indicate that,
for the Ricker, Hassell, and Pennycuick models, left-tail association accentuates and
right-tail association mitigates extinction risk for undercompensatory dynamics ($r<r_c$),
whereas the reverse occurs for overcompensatory dynamics ($r>r_c$). Fig. \ref{fig_extrisk_d0}
reveals a similar reversal for the Maynard Smith, Verhulst and Austin-Brewer models; but 
it occurs at a value of $r$ close to but distinguishable from $r_c$. Inspection of Figs
\ref{SM-fig_ricker_risk_local_disp_scl_1}-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
shows that a qualitatively similar conclusion holds for nonzero $d$, for both local and global
dispersal.

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (C) \hspace{4 cm} (D)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}
\caption{Extinction risk versus time for our stochastic, age-structured model. Risks
were higher when associations between vital rates were stronger in the left compared to
the right tails of the rates, but only slightly so. Results are for the linear
version of the model ($\Omega = \infty$; A, C) and for the nonlinear model with density
dependence in the fecundity rate ($\Omega = 100$; B, D). Results show cases of moderate
(A, B) and extreme (C, D) asymmetry of tail association. The solid line is the estimated extinction risk, and the area of shading, which is barely visible because it is very narrow, shows 95\% confidence intervals for extinction risk based on the number of simulations done. LT=left tail, RT=right tail.\label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of the six density-dependent stochastic metapopulation models we used. The population growth rate in habitat patch $i$ at time $t$, $\\lambda_i(t)$ (second column), is $P_i(t+1)/P_i(t)$, and defines how the model works within each patch. Here $P_i(t)$ is the population in patch $i$ at time $t$. The growth rate depends on model parameters, $P_i(t)$, and the environment $\\varepsilon_i(t)$ in patch $i$ at time $t$. $\\varepsilon_i(t)$ was normally distributed with mean $0$ and standard deviation $\\sigma$, and values were generated independently across times. $P_e$ is the equilibrium of the deterministic, one-patch version of the model. $r_{min}$ is the smallest $r$, given fixed values of the other parameters, for which $P_e$ was non-negative. $r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text). $r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilibrium (a bifurcation; see main text). For extinction-risk simulations, we used $r$ such that $r_{min}<r<r_{bf}$. Dispersal between patches was also implemented (see main text). Formulas are given for $P_e$ but specific values which pertain for the listed parameters are given for $r_{min}$, $r_c$ and $r_{bf}$. See Appendix \\ref{SM-model_stability} for the formulas that gave these values.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6,7),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```

>>>>

<!--stochastic matrix model : age-structured-->
To investigate Q1 from the Introduction,
we considered a simple stochastic age-structured model
with two age classes, 
\begin{equation}
\left({\begin{array}{cc} E(t+1) \\ A(t+1) \end{array}}\right) = 
\left(\begin{array}{cc} 0 & f_a(t)\\ s_e(t) & 0 \end{array}\right)
\left(\begin{array}{c} E(t) \\ A(t) \end{array}\right), \label{eq_age_str}
\end{equation}
\noindent where $E(t)$ and $A(t)$ are the numbers or population densities of eggs 
and adults, respectively, at time $t$. We chose a very simple model
because our goal was to carry out an initial evaluation, only, of 
the realism of a potential mechanism. The stochastic vital 
rates $s_e(t)$ and $f_a(t)$ in Eq. \ref{eq_age_str} were selected, independently 
across times, as follows. The egg survival rate $s_e(t)$ was 
drawn from a beta distribution with parameters $\alpha=3$ and $\beta=2$. 
Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
was drawn from a gamma distribution with shape and scale parameters $k=2$
and $\theta=1$, respectively, 
and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$
$\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
(it was then a standard stochastic matrix model). Finite, positive values of 
$\Omega$ made the model nonlinear, with density dependence in the adult
fecundity rate. Stochastic matrix models are very commonly used in
evolutionary ecology and quantitative conservation biology, so our model
is an example of a very widely used category of models. See, for instance,
the classic books of @Caswell2000, and @MorrisDoak2002.

@@

The variables $u(t)$ and $v(t)$ are visibly asymmetrically tail associated (Fig. \ref{fig_intro_cop}),
but we have instead used the transformed variables $s_e(t)$ and $\tilde{f}_a(t)$
in our model. In what sense are $s_e(t)$ and $\tilde{f}_a(t)$ asymmetrically
tail associated? Given two boundary quantities $l_b$ and $u_b$ such
that $0 \leq l_b < u_b \leq 1$, and letting $x(t)$ be the rank of 
$u(t)$ in the set $\{u(t) | t=1,\ldots,T\}$, divided by $T+1$, and letting 
$y(t)$ be the rank of 
$v(t)$ in the set $\{v(t) | t=1,\ldots,T\}$, divided by $T+1$ (smallest elements
are here defined to have rank 1),
consider the portions of the 
lines $x+y = 2l_b$ and $x+y=2u_b$ which lie within the unit square
(Fig. \ref{PartialSpearmanFig}). 
We can then define the *partial Spearman correlation*,
\begin{equation}
\cor_{l_b,u_b}(u,v) = \frac{\sum (x(t)-\mean(x))(y(t)-\mean(y))}{(T-1)\sqrt{\var(x) \var(y)}},\label{PartialSpearmanEq}
\end{equation}
\noindent where sample means and variances in this equation
are computed using all the data but the sum is computed only
over the $t$ such that $x(t)+y(t) > 2l_b$ and $x(t)+y(t)<2u_b$. The variance, $\var$,
here and throughout, is the sample variance, with Bessel's correction (denominator $T-1$).
The partial Spearman correlation is the component of the Spearman
correlation of $u$ and $v$ which can be attributed to the points
for which the normalized ranks, $x$ and $y$, lie between the boundary lines
we defined in
the unit square (Fig. \ref{PartialSpearmanFig}). 
The quantity $\cor_{0,0.5}(u,v)-\cor_{0.5,1}(u,v)$ 
is a way to measure asymmetry of tail associations between
$u$ and $v$. And this quantity is the same as 
$\cor_{0,0.5}(s_e,\tilde{f}_a)-\cor_{0.5,1}(s_e,\tilde{f}_a)$, 
because the partial Spearman correlation, like the Spearman
correlation itself, is based on ranks. Ranks are the same
for $u(t)$ and for $s_e(t)$ because these quantities are 
related by an increasing monotonic transformation; and likewise for 
$v(t)$ and $\tilde{f}_a(t)$. Thus, in the sense captured
by the statistic $\cor_{0,0.5}-\cor_{0.5,1}$, $s_e(t)$
and $\tilde{f}_a(t)$ have the same asymmetry of tail
association as the values $u(t)$ and $v(t)$ from which they
were constructed. Rank-based approaches to understanding the nature
of association between variables have been recommended, as such
approaches are not influenced by the marginal distributions of
the quantities (e.g., @Genest2007). Marginal distributions contain no information
about association between variables, as each marginal pertains to
only one of the variables. These ideas are revisited in the Discussion.

@@

<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
To answer (Q2) of the Introduction, 
we estimated extinction risk for each metapopulation model, using each form of tail 
association for the environmental noise,
using values of $r$ that included under- and overcompensatory dynamics,
as indicated above. We chose values of $r$ within a range, the
range specified as follows. The second part of Q2 (about how density dependence mediates the
influence of asymmetric tail associations on extinction risk) was addressed by comparing
results for under- and overcompensatory model regimes.
Defining $r_\text{min}$ as the minimum value of $r$ 
for which $P_e$ was non-negative; and defining
$r_\text{bf}$ as the value of $r$ at which a bifurcation occurs, the range of
$r$ used spanned 95\% of the interval ($r_\text{min},r_\text{bf}$). For details, see Appendix 
\ref{SM-model_stability}. The range of $r$ encompassed $r_c$, so included 
a transition from under- to overcompensatory dynamics. For each value of $r$
used within the range, we calculated extinction risk as indicated above.
Dispersal parameter values $d= 0,0.1,0.2,\ldots,0.5$ were also used in different 
simulations. We plotted extinction risk as a function of $d$ for several
fixed values of $r$, some producing under- and some overcompensatory dynamics.
For $d=0$ we also plotted extinction risk as a function of $r$. All code for 
this project is available at https://github.com/sghosh89/ERC.

@@

<!--Q2 : metapopulation results-->
For our metapopulation simulations, for low values of $r$ (close to $r_\text{min}$), left-tail
associations of environmental noises accentuated extinction risks, and right-tail associations
mitigated them; but the reverse was true for high values of $r$ (close to
$r_\text{bf}$). Thus 
density dependent models can either produce similar results, regarding the influence of 
tail associations on extinction risk, as the density-independent models of @Ghosh_copula, or
opposite results, depending on the nature of density dependence. 
For each model of Table \ref{tab_model_summary}, using a value of $r$ on
the low end of the range $(r_\text{min},r_\text{bf})$, extinction risk was highest
for extreme left-tail associated noise, lowest for extreme right-tail associated
noise, and was intermediate for the moderate tail association scenarios, for
both local (Fig. \ref{fig_extrisk_allmodel_localdisp}, left panels) and global 
(Fig. \ref{SM-fig_extrisk_allmodel_globaldisp}, left panels) dispersal, and for 
essentially all the dispersal rates, $d$, that we considered. However, 
using a value of $r$ on the high end of the range $(r_\text{min},r_\text{bf})$,
extinction risks were ordered oppositely (right panels of Figs 
\ref{fig_extrisk_allmodel_localdisp} and \ref{SM-fig_extrisk_allmodel_globaldisp}). 
These results help address Q2 from the Introduction.
Intermediate values of $r$ are shown in Figs. \ref{SM-fig_ricker_risk_local_disp_scl_1} - 
\ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} for local dispersal and 
Figs. \ref{SM-fig_ricker_risk_global_disp_scl_1} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6} for global dispersal among 
habitat patches.

Because the deterministic, one-patch versions of our models showed undercompensatory
dynamics for $r<r_c$ (and therefore for $r$ on the left end of the range $(r_\text{min},r_\text{bf})$),
and showed overcompensatory dynamics for $r>r_c$, Figs \ref{fig_extrisk_allmodel_localdisp}
and \ref{SM-fig_extrisk_allmodel_globaldisp} suggest the hypothesis that the
opposite effects of tail association on extinction risk correspond to the domains
in which models exhibit under- versus overcompensatory dynamics. This was 
approximately, but not exactly, true. Careful inspection of Figs 
\ref{SM-fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{SM-fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6} 
and the $r_c$ values of Table \ref{tab_model_summary} reveals that, for the Verhulst and Austin-Brewer models,
the switch occurs at a value of $r$ less than $r_c$. For the example
case of $d=0$, plots of extinction risk versus $r$ (Fig. \ref{fig_extrisk_d0}) seem to indicate that,
for the Ricker, Hassell, and Pennycuick models, left-tail association accentuates and
right-tail association mitigates extinction risk for undercompensatory dynamics ($r<r_c$),
whereas the reverse occurs for overcompensatory dynamics ($r>r_c$). Fig. \ref{fig_extrisk_d0}
reveals a similar reversal for the Maynard Smith, Verhulst and Austin-Brewer models; but 
it occurs at a value of $r$ close to but distinguishable from $r_c$. Inspection of Figs
\ref{SM-fig_ricker_risk_local_disp_scl_1}-\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
shows that a qualitatively similar conclusion holds for nonzero $d$, for both local and global
dispersal.

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (C) \hspace{4 cm} (D)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}
\caption{Extinction risk versus time for our stochastic, age-structured model. Risks
were higher when associations between vital rates were stronger in the left compared to
the right tails of the rates, but only slightly so. Results are for the linear
version of the model ($\Omega \rightarrow \infty$; A, C) and for the nonlinear model with density
dependence in the fecundity rate ($\Omega = 100$; B, D). Results show cases of moderate
(A, B) and extreme (C, D) asymmetry of tail association. The solid line is the estimated extinction risk, and the area of shading, which is barely visible because it is very narrow, shows 95\% confidence intervals for extinction risk based on the number of simulations done. LT=left tail, RT=right tail.\label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of the six density-dependent stochastic metapopulation models we used. The population growth rate in habitat patch $i$ at time $t$, $\\lambda_i(t)$ (second column), is $P_i(t+1)/P_i(t)$, and defines how the model works within each patch. Here $P_i(t)$ is the population in patch $i$ at time $t$. The growth rate depends on model parameters, $P_i(t)$, and the environment $\\varepsilon_i(t)$ in patch $i$ at time $t$. $\\varepsilon_i(t)$ was normally distributed with mean $0$ and standard deviation $\\sigma$, and values were generated independently across times. $P_e$ is the equilibrium of the deterministic, one-patch version of the model. $r_\\text{min}$ is the smallest $r$, given fixed values of the other parameters, for which $P_e$ was non-negative. $r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text). $r_\\text{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilibrium (a bifurcation; see main text). For extinction-risk simulations, we used $r$ such that $r_\\text{min}<r<r_\\text{bf}$. Dispersal between patches was also implemented (see main text). Formulas are given for $P_e$ but specific values which pertain for the listed parameters are given for $r_\\text{min}$, $r_c$ and $r_\\text{bf}$. See Appendix \\ref{SM-model_stability} for the formulas that gave these values.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6,7),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```
